---
title: "Exploratory Analysis"
output: 
  html_document:
    toc: true
    toc_float: true
---
```{r setup, include = FALSE}
library (tidyverse)
library (readxl)
library(plotly)


knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

vaccine_coverage_df = read_csv(file = "./data/coverage_by_modzcta_adults.csv") %>% 
  janitor::clean_names()

```

### **1. The vaccination coverage in NYC**

MADE SOME SCATTERPLOTS with smooth line

```{r plot1}
ggplot(vaccine_coverage_df,aes (x = pop_denominator, y = count_fully_cumulative, color = borough)) + geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ borough)

ggplot(vaccine_coverage_df,aes (x = pop_denominator, y = count_partially_cumulative, color = borough)) + geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ borough)

```

MADE SOME SIMPLE SCATTERPLOTS AND THEIR GGPLOTLY 

```{r}


scatter_ggplot1 = 
vaccine_coverage_df   %>%
  ggplot(aes(x = pop_denominator, y = count_partially_cumulative, color = borough)) +
  geom_point(alpha = 1) +
  coord_cartesian()

ggplotly(scatter_ggplot1)

ggplot(vaccine_coverage_df,aes (x = pop_denominator, y = count_partially_cumulative, color = borough)) + geom_point()


ggplot(vaccine_coverage_df,aes (x = count_fully_cumulative, fill = borough)) + geom_histogram()

ggplot(vaccine_coverage_df,aes (x = count_partially_cumulative, fill = borough)) + geom_histogram()

```

MADE SOME DENSITY PLOTS WITH A DENISTY BAR BELOW

```{r}
ggplot(vaccine_coverage_df, aes(x = count_fully_cumulative, fill = borough)) + 
  geom_density(alpha = .4, adjust = .5, color = "blue") + geom_rug()

ggplot(vaccine_coverage_df, aes(x = count_partially_cumulative, fill = borough)) + 
  geom_density(alpha = .4, adjust = .5, color = "blue") + geom_rug()

```

### **2. The booster coverage in NYC**

```{r}
ggplot(vaccine_coverage_df,aes (x = pop_denominator, y = count_additional_cumulative, color = borough)) + geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ borough)

ggplot(vaccine_coverage_df,aes (x =pop_denominator , y = count_additional_cumulative, color = borough)) + geom_point()

ggplot(vaccine_coverage_df,aes (x = count_additional_cumulative, fill = borough)) + geom_histogram()

ggplot(vaccine_coverage_df, aes(x = count_additional_cumulative, fill = borough)) + 
  geom_density(alpha = .4, adjust = .5, color = "blue")+ geom_rug()

```

### **3. Vaccination and booster coverage vs The Zip Code data for NYC**


### **4. The Education Attainment vs vaccination and booster coverage **

### **5. The Poverty Status vs vaccination and booster coverage**

```{r include = FALSE}
poverty_status_df = read_csv(file = "./data/poverty_status.csv", skip = 1) %>% 
  janitor::clean_names()

poverty_status_tidy = poverty_status_df%>% 
  mutate (modzcta = substr(geographic_area_name,6,11),total_pop_pov_above_18 = estimate_total_population_for_whom_poverty_status_is_determined - estimate_total_population_for_whom_poverty_status_is_determined_age_under_18_years) %>% 
  select(modzcta, estimate_total_population_for_whom_poverty_status_is_determined, total_pop_pov_above_18)
```


```{r show_col_types = FALSE, ECHO = FALSE, include = FALSE}
zipcode_df = read_csv(file = "./data/Modified_Zip_Code_Tabulation_Areas__MODZCTA_.csv")
```

```{r vaccine-data, include = FALSE}
vaccine_data = read.csv(file = "./data/coverage_by_modzcta_adults.csv") %>%
  janitor::clean_names() %>%
  mutate(modzcta = as.character(modzcta))

vaccine_data$modzcta <- gsub('\\s+', '', vaccine_data$modzcta)
poverty_status_tidy$modzcta <- gsub('\\s+', '', poverty_status_tidy$modzcta)
```

```{r merge-dataset, include = FALSE}
joined_dataset_poverty = merge(vaccine_data,poverty_status_tidy, by = "modzcta", all.x = TRUE)
```

#### total est population poverty status above 18 vs total pop by borough
```{r}
ggplot(joined_dataset_poverty, aes ( x=total_pop_pov_above_18, y = estimate_total_population_for_whom_poverty_status_is_determined, color = borough)) + geom_point()
```

#### Total pop pov above 18 vs fully vaccinated by borough
```{r}
ggplot(joined_dataset_poverty, aes (x = total_pop_pov_above_18, y = count_fully_cumulative, color = borough)) + geom_point()
```

#### Total pop pov above 18 vs partially vaccinated by borough
```{r}
ggplot(joined_dataset_poverty, aes (x = total_pop_pov_above_18, y = count_fully_cumulative, color = borough)) + geom_point()
```



### **6. The Median Income vs vaccination and booster coverage**





