---
title: "Exploratory Analysis"
output: 
  html_document:
    toc: true
    toc_float: true
---
```{r setup, include = FALSE}
library (tidyverse)
library (readxl)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

vaccine_coverage_df = read_csv(file = "./data/coverage_by_modzcta_adults.csv")

```

### **1. The vaccination coverage in NYC**

MADE SOME SCATTERPLOTS with smooth line

```{r plot1}
ggplot(vaccine_coverage_df,aes (x = POP_DENOMINATOR, y = COUNT_FULLY_CUMULATIVE, color = BOROUGH)) + geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ BOROUGH)

ggplot(vaccine_coverage_df,aes (x = POP_DENOMINATOR, y = COUNT_PARTIALLY_CUMULATIVE, color = BOROUGH)) + geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ BOROUGH)

ggplot(vaccine_coverage_df,aes (x = POP_DENOMINATOR, y = COUNT_ADDITIONAL_CUMULATIVE, color = BOROUGH)) + geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ BOROUGH)
```

MADE SOME SIMPL SCATTERPLOTS

```{r}
ggplot(vaccine_coverage_df,aes (x = POP_DENOMINATOR, y = COUNT_FULLY_CUMULATIVE, color = BOROUGH)) + geom_point()

ggplot(vaccine_coverage_df,aes (x = POP_DENOMINATOR, y = COUNT_PARTIALLY_CUMULATIVE, color = BOROUGH)) + geom_point()

ggplot(vaccine_coverage_df,aes (x = POP_DENOMINATOR, y = COUNT_ADDITIONAL_CUMULATIVE, color = BOROUGH)) + geom_point()

ggplot(vaccine_coverage_df,aes (x = COUNT_FULLY_CUMULATIVE, fill = BOROUGH)) + geom_histogram()

ggplot(vaccine_coverage_df,aes (x = COUNT_PARTIALLY_CUMULATIVE, fill = BOROUGH)) + geom_histogram()

ggplot(vaccine_coverage_df,aes (x = COUNT_ADDITIONAL_CUMULATIVE, fill = BOROUGH)) + geom_histogram()

```

MADE SOME DENSITY PLOTS WITH A DENISTY BAR BELOW

```{r}
ggplot(vaccine_coverage_df, aes(x = COUNT_FULLY_CUMULATIVE, fill = BOROUGH)) + 
  geom_density(alpha = .4, adjust = .5, color = "blue") + geom_rug()

ggplot(vaccine_coverage_df, aes(x = COUNT_PARTIALLY_CUMULATIVE, fill = BOROUGH)) + 
  geom_density(alpha = .4, adjust = .5, color = "blue") + geom_rug()

ggplot(vaccine_coverage_df, aes(x = COUNT_ADDITIONAL_CUMULATIVE, fill = BOROUGH)) + 
  geom_density(alpha = .4, adjust = .5, color = "blue")+ geom_rug()


```


### **2. The booster coverage in NYC**

### **3. Vaccination and booster coverage vs The Zip Code data for NYC**

### **4. The Education Attainment vs vaccination and booster coverage **

### **5. The Poverty Status vs vaccination and booster coverage**

### **6. The Median Income vs vaccination and booster coverage**
